A Particle Swarm Optimisation-Based Indirect Approach to Web Service Composition

### 2015-10-09 paper.tex
Finished writing layered mapping section; added notes for new fitness calculation.

A wide variety of Evolutionary Computing (EC) approaches has been applied to the problem of Web service composition, as evidenced by literature surveys of the field \cite{wang2012survey,pejman2012web}. One of the earliest works in this area \cite{canfora2005approach} applies genetic algorithms to optimise the overall Quality of Service (QoS) of a composition. The composition process in this work is \textit{semi-automated}, meaning that a workflow of abstract services has already been provided. In semi-automated compositions, the objective is to select a set of concrete services that fulfil the required functionality of their abstract counterparts, ensuring that the selected set results in a composition with the best possible quality. Even this approach takes QoS into account, it is not capable of performing \textit{fully automated} composition, which is when the composition workflow is automatically deduced at the same time that the services to include in the composition are identified.

* A semi-automated PSO approach to Web service composition (\cite{zhao2012improved})

@@

* QoS-aware Web GP approach (use Yang's \cite{yu2013adaptive})

* GraphEvol approach \cite{da2015graphevol}

@@

>>>>

A wide variety of Evolutionary Computing (EC) approaches has been applied to the problem of Web service composition, as evidenced by literature surveys of the field \cite{wang2012survey,pejman2012web}. One of the earliest works in this area \cite{canfora2005approach} applies genetic algorithms to optimise the overall Quality of Service (QoS) of a composition. The composition process in this work is \textit{semi-automated}, meaning that a workflow of abstract services has already been provided. In semi-automated compositions, the objective is to select a set of concrete services that fulfil the required functionality of their abstract counterparts, ensuring that the selected set results in a composition with the best possible quality. Even though this approach takes QoS into account, it is not capable of performing \textit{fully automated} composition, which is when the composition workflow is automatically deduced at the same time that the services to include in the composition are identified. Several works employ particle swarm optimisation (PSO) to solving the problem of service composition \cite{wang2010improved,amiri2012effective,ludwig2012applying,zhao2012improved}, but similarly to genetic algorithm they focus exclusively on semi-automated composition.

@@

Finally, some approaches both create the composition workflow and optimise the quality of the overall composition \cite{yu2013adaptive,da2015graphevol}. These works accomplish this by relying on variable-size solution representations (trees or directed acyclic graphs) and by measuring the quality of candidate compositions through the fitness function. In \cite{yu2013adaptive}, the fitness function is responsible for penalising solutions that are not \textit{functionally correct}, i.e. solutions that contain services whose inputs have not been entirely fulfilled; in \cite{da2015graphevol}, candidate initialisation and genetic operators are restricted to only produce functionally correct solutions. While these approaches do consider both workflow creation and quality improvement simultaneously, they do not consider the idea of optimising solutions indirectly, and instead perform operations to the solution workflows directly.

@@

\section{Layered PSO-based Composition Approach}

A variation to the PSO-based approach proposed in the previous section, which considers the use of service layers, was also developed. As shown in Algorithm \ref{novelLayeredSteps}, this approach follows the same steps used in the non-layered method, with three fundamental differences: firstly, the mapping of candidates to an index in the particle's position vector includes the division of the services that can potentially be included into a composition for the given task into layers; secondly, the the decoding and evaluation of the solution represented by a particle are performed in a single step by using a new algorithm that does not require the creation of a graph; finally, a graph representing the global best solution is created after the optimisation process has finished running. Each of these steps is explained in the subsections below.

\begin{algorithm}[h]
 \setlength\hsize{0.9\linewidth}
 \SetKwInOut{Input}{Input}\SetKwInOut{Output}{Output}
 \let\oldnl\nl% Store \nl in \oldnl
\newcommand{\nonl}{\renewcommand{\nl}{\let\nl\oldnl}}
 \LinesNumbered
 	\textcolor{blue}{\textbf{1.} Identify layers and map each candidate service to an index in the particle's position vector.}\\
	\textbf{2.} Randomly initialise each particle in the swarm.\\
	\nonl \While {max. iterations not met}{
		\ForAll{particles in the swarm}{
		    \textcolor{blue}{\textbf{3.} Decode and calculate quality of corresponding solution layer by layer.}\\
			\eIf{fitness value better than pBest}{
				\textbf{4a.} Assign current fitness as new pBest.\\
			}{
				\textbf{4b.} Keep previous pBest.\\
			}		
		}
		\textbf{5.} Assign best particle's pBest value to gBest, if better than gBest.\\
		\textbf{6.} Calculate the velocity of each particle according to the equation:\\
		\Indp $v_{id} = v_{id} + c_1 * rand() * (p_{id} - x_{id}) + c_2 * rand() * (p_{gd} - x_{id})$\\
		\Indm \textbf{7.} Update the position of each particle according to the equation:\\
		\Indp $x_{id} = x_{id} + v_{id}$\\
		\Indm \textcolor{blue}{\textbf{8.} Create final solution graph based using gBest's weights.}\\
	}

 \caption{Steps of the layered PSO-based Web service composition technique.}
\label{novelLayeredSteps}
\end{algorithm}

\subsection{Layer Identification and Particle Mapping (Step 1)}

The unique feature of the PSO approach proposed in this section is that it identifies the \textit{composition layer} to which a service belongs. The layers are determined during the discovery process that identifies the services that could be possibly used in the composition, which means that Algorithm \ref{discovery} is revised to keep track of composition layers, as shown in the extended Algorithm \ref{layeredDiscovery}. Each time a new set $S_{found}$ of services is discovered, it is classified as a new composition layer. The implication of this classification is that the inputs of a service in a given layer can be completely satisfied by using the services in previously discovered layers, which later facilitates the solution decoding and fitness calculation process.

\begin{algorithm}
\setlength\hsize{0.9\linewidth}
 \SetKwInOut{Input}{Input}\SetKwInOut{Output}{Output}
 \LinesNumbered
 \SetNlSty{}{}{:}
 \Input{$I$, $O$, $R$}
 \Output{a list of services $S_{list}$}
 $C_{search} \leftarrow I$\;
 $S_{list} \leftarrow \{\}$\;
 $S_{found} \leftarrow DiscoverService()$\;
 \While{$|S_{found}| > 0$}{
  $S_{list} \leftarrow S_{list} \cup S_{found}$\;
  $C_{search} \leftarrow C_{search} \cup C_{output}$ of $S_{found}$\;
  $S_{found} \leftarrow DiscoverService()$\;
  }
  \eIf{$C_{search} \supseteq O$}{
   \Return $S_{list}$\;
   }{
   Report no solution\;
 }
\caption{Discovering relevant services for compositio and tracking composition layers.}
\label{layeredDiscovery}
\end{algorithm}

Just as before, services are mapped to particle indices once the discovery process has taken place. This time, however, the mapping is organised so that

\subsection{Solution Decoding and Fitness Calculation (Step 3)}

\subsection{Construction of Final Graph (Step 8)}
